---
title: "Draft_porous"
author: "Lorenzo Menichetti"
date: "2023-02-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(SoilR)
```


Here the model draft. I did not implement yet the nonlinearity at the input level yet, but added the bioturbation feedback fluxes $T_y=k_{mix} \frac{M_{y_{(mic)}}-M_{y_{(mes)}}}{2}$ and $T_o=k_{mix} \frac{M_{o_{(mic)}}-M_{0_{(mes)}}}{2}$.
I consider them as a linear flux, no mistakes here, right?  

My question now is: how can I implement some rather complicated function for the input nonlinearity?  
I would think to build two functions $phi_{mes}=f(M_{y_{(mes)}}, M_{o_{(mes)}}, M_{y_{(mic)}}, M_{o_{(mic)}})$ and $phi_{mic}=f(M_{y_{(mic)}}, M_{o_{(mic)}})$, but how to call them inside the model?

```{r}

Porous<-function(ky=0.8, ko=0.00605,
                 kmix=0.9,
                 e=0.13,
                 Im=1.1, Ir=0.5,
                 F_prot=0.0){
  time_symbol='t'

  ##### IN
  ifs=SoilR:::InFluxList_by_PoolName(
    c(
      SoilR:::InFlux_by_PoolName(
        destinationName='My_mes',
        func=function(t){
          Im
        }
      ),
      SoilR:::InFlux_by_PoolName(
        destinationName='My_mes',
        func=function(t){
          Ir*(0.5)
        }
      ),
      SoilR:::InFlux_by_PoolName(
        destinationName='My_mic',
        func=function(t){
          Ir*(0.5)
        }
      )
    )
  )
  ##### OUT
  ofs=SoilR:::OutFluxList_by_PoolName(
    c(
      SoilR:::OutFlux_by_PoolName(
        sourceName='My_mes',
        func=function(My_mes){
          ky*My_mes
        }
      )
      ,
      SoilR:::OutFlux_by_PoolName(
        sourceName='Mo_mes',
        func=function(Mo_mes){
          (1-e)*ko*Mo_mes
        }
      ),
      SoilR:::OutFlux_by_PoolName(
        sourceName='My_mic',
        func=function(My_mic){
          ky*F_prot*My_mic
        }
      ),
      SoilR:::OutFlux_by_PoolName(
        sourceName='Mo_mic',
        func=function(Mo_mic){
          (1-e)*ko*F_prot*Mo_mic
        }
      ),
      # Bioturbation, nonlinearity here, Ty and To are outfluxes from the micropores (and also internal fluxes below, into My_mes and Mo_mes)
      #Ty
      SoilR:::OutFlux_by_PoolName(
        sourceName='My_mic',
        func=function(My_mic,My_mes){
          kmix*((My_mic-My_mes)/2)
        }
      ),
      #To
      SoilR:::OutFlux_by_PoolName(
        sourceName='Mo_mic',
        func=function(Mo_mic,Mo_mes){
          kmix*((Mo_mic-Mo_mes)/2)
        }
      )
    )
  )

  ##### INT
  intfs=SoilR:::InternalFluxList_by_PoolName(
    list(
      SoilR:::InternalFlux_by_PoolName(
        sourceName='My_mes',
        destinationName='Mo_mes',
        func=function(My_mes){
          e*ky*My_mes
        }
      ),
      SoilR:::InternalFlux_by_PoolName(
        sourceName='My_mic',
        destinationName='Mo_mic',
        func=function(My_mic){
          e*ky*F_prot*My_mic
        }
      ),
      #bioturbation fluxes
      #Ty
      SoilR:::InternalFlux_by_PoolName(
        sourceName='My_mic',
        destinationName='My_mes',
        func=function(My_mic, My_mes){
          kmix*((My_mic-My_mes)/2)
        }
      ),
      #To
      SoilR:::InternalFlux_by_PoolName(
        sourceName='Mo_mic',
        destinationName='Mo_mes',
        func=function(Mo_mic, Mo_mes){
          kmix*((Mo_mic-Mo_mes)/2)
        }
      )
    )
  )

  smod <- SoilR:::SymbolicModel_by_PoolNames(
    in_fluxes=ifs,
    internal_fluxes=intfs,
    out_fluxes=ofs,
    timeSymbol=time_symbol
  )
  smod
}
```



```{r}
modelObject<-Porous()
plotPoolGraph(modelObject)
```

Here I have some doubts... the graph seems inconsistent, even if the syntax seems according to your specs. From $M_{y_{mic}}$ and $M_{o_{mic}}$ we have three outfluxes represented in the graph, but they should be just two. I do not think the error below is because of that, but still it looks weird.  
If it works it's fine of course, is  it just an inconsistency in the graphing?  
But it runs and seems reasonable.

```{r}
iv<-c(My_mes=1, Mo_mes=10,My_mic=0.6, Mo_mic=3)
times<-seq(0,20,by=0.1)
modrun0<-Model_by_PoolNames(smod=modelObject, times=times, initialValues=iv)
```

```{r, error=TRUE}
Ct0<-getC(modrun0)
matplot(times, Ct0, type="l", lty=1, col=1:4, xlab=" ", ylab="Pool contents", bty="n")
legend("topleft", c("My_mes", "Mo_mes", "My_mic", "Mo_mic"), lty=1, col=1:4, bty="n")

```
